### 2017-02-02 CO750-ApproximationAlgorithmsTopics/assignments/assignment_01.tex
Retyped 3b of apx algs

Our goal is to find a maximum $k$-Cut of $G$. That is, we seek to find a partition of $V$ into $V_1, \dots, V_k$ maximizing
$$\sum_{i=1}^k \sum_{j = 1}^{i-1} |E(V_i, V_j)| := c(\{V_1, \dots, V_k\}).$$
Intuitively we are maximizing the edges between the selected cuts. Let $\cB$ denote the $2$-approximation algorithm for max cut given in problem $3a$. Consider the following algorithm, which we will denote by $\cA$:

@@

\item Set $V_1 = S$ and $V_2 = V\backslash S$ where $S$ is the cut returned by $\cB$ when run on $G$.
\item Set $V_i = \emptyset$ for $i = 3, \dots, k$.
\item For $i = 2, \dots, k-1$

@@

\item For $j = 1, \dots, i$ set $S_j$ to the cut returned by $\cB$ when run on $G[V_j]$.
\item Set $j^* = \text{argmax}_{j=1,\dots, i} \{|E(S_j, V_j\backslash S_j)|\}$.
\item Set $V_{i+1} = S_{j^*}$ and set $V_{j^*} = V_{j^*} \backslash S_{j^*}$.

@@

The idea behind the operation of $\cA$ is to start with a $2$-cut of $G$, and in every successive iteration take the current partition and $2$-cut one of the parts in an approximately optimal way until we reach a $k$-cut.

@@

Let $\cO = \{V^*_1, \dots, V^*_k\}$. Consider an arbitrary iteration $i = 1, \dots, k-1$ (iteration $1$ is original $2$-cut). Let $\cS = \{V_1, \dots, V_{i+1}\}$ be the set of partitions chosen so far by $\cA$. Suppose that at the end of iteration $i$, the algorithm has split $V_j$ into $V_j'$ and $V_{i+1}$. That is the solution after iteration $i$ is
$$\cS' = \cS \backslash V_j \cup \{V_j', V_{i+1}\}.$$
$$2k(c(\cS') - c(\cS)) \geq c(\cO) - c(\cS).$$

@@

The difference between $c(\cS')$ and $c(\cS)$ is precisely $|E(V_j', V_{i+1})|$ (the number of edges gained by splitting $V_j$). Further the difference $c(\cO) - c(\cS)$ is upper-bounded by the edges counted in $c(\cO)$ that are not counted in $c(\cS)$. Call such edges $E'$.  The edges $E'$ are necessarily internal to a part of $\cS$ (that is, both endpoints lie in some $V_k \in \cS$). Otherwise they are counted in $c(\cS)$ since their endpoints lie in distinct parts of $\cS$. Thus if we consider the $k$-cut induced by $\cO$ on $G[E']$ it consists of at most $k$ splits ($2$-cuts) of parts $V_k \in \cS$.
Let $V'_k$ denote the best possible $2$-cut of $V_k \in \cS$ for all $k = 1, \dots, i$. Now let $$k^* = \text{argmax}_{k=1,\dots,i}  \{|E(V'_k, V_k \backslash V'_k)|\}.$$ We have
$$ |E'| \leq k |E(V'_{k^*}, V_k \backslash V'_{k^*})|$$
by our previous discussion, since $E'$ edges are formed by a $k$-cut on edges internal to parts of $\cS$, and  $|E(V'_{k^*}, V_k \backslash V'_{k^*}|$ is largest $2$-cut to be formed by edges internal to a part of $\cS$. Now the cut found by our algorithm $\cA$ in this iteration in step $3(c)$ is a $2$-approximation of the best possible split cut. That is,
$$|E(V'_{k^*}, V_k \backslash V'_{k^*})| \leq 2|E(V_j', V_{i+1})| .$$
Hence, combining these inequalities, we see that
$$|E'| \leq 2k |E(V_j', V_{i+1})|.$$
Therefore $2k(c(\cS') - c(\cS)) \geq c(\cO) - c(\cS)$ as desired.

@@

Since $\cB$ can be run in polynomial time, and $k$ is at most $|V|$, it is clear that $\cA$ runs in polynomial time. At termination the algorithm returns a partition of $V$ into $k$ parts so it returns a feasible solution.

@@

Now to see the approximation factor holds let $\cS^i$ denote the solution maintained by $\cA$ at the end of iteration $i$. Then the returned solution is $\cS^{k-1}$. Let $\cO$ be an optimal solution. Observe from Lemma \ref{lemma:3b1} that $c(\cS^{i}) \geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{i-1})$. We will apply this inequality inductively to achieve the bound:

@@

c(\cS^{k-1} &\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{k-2}) \\
		&\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})(\frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{k-3})) \\

@@

		&= 1-(1-\frac{1}{2k})^kc(\cO).

>>>>

Our goal is to find $k$ cuts of $G$ $V_1, \dots, V_k \subseteq V$ maximizing:
$$|\bigcup_{i=1}^k \delta(V_i)|.$$
Let $\cB$ denote the $2$-approximation algorithm for max cut given in problem $3a$. Consider the following algorithm, which we will denote by $\cA$:

@@

\item Set $V_1 = \dots = V_k = \emptyset$. Set $G_1 = G$
\item For $i = 1, \dots, k$

@@

\item Let $S$ be the cut returned by $\cB$ run on $G_i$.
\item Set $V_i = S$. Set $G_{i+1} = G_i \backslash \delta(S).$

@@

The idea behind the operation of $\cA$ is to find an optimal cut, then remove the edges of the cut and iterate.

@@

Let $\cO = \{V^*_1, \dots, V^*_k\}$ denote the optimal solution. For any $i=1,\dots, k$ let $\cS^i = \{V_1, \dots, V_{i}\}$ be the set of partitions chosen so far at the end of iteration $i$ by $\cA$. Then 
$$2k(c(\cS^i) - c(\cS^{i-1})) \geq c(\cO) - c(\cS^{i-1}).$$
With $S^0 = \emptyset$.

@@

For $i = 1,\dots, k$ let $S_i^* = V_i^k \backslash \bigcup_{S\in \cS^{i-1}} S$. Then
$$c(\cO) - c(\cS^{i-1}) = c(\{S^*_1,\dots, S^*_l\}) = |\bigcup_{i=1}^k \delta(S^*_i)|.$$
Let $$S^* = \text{arg max}_{S^*_i : i=1,\dots, k} |\delta(S^*_i)|.$$
$$c(\cO) - c(\cS^{i-1}) \leq k|\delta(S^*)|.$$
Now observe that for all $i = 1,\dots, k$, $\delta(S^*_i) \subseteq E(G_i)$, hence in particular $\delta(S^*) \subseteq E(G_i)$. Thus $|\delta(S^*)|$ has size at most that of an optimal cut in $G_i$. Now by our choice of $V_i$, using $2$-approximation algorithm $\cB$, $2|\delta(V_i)|$ is at the size of an optimal cut in $G_i$. Hence we have
$$|\delta(S^*)| \leq 2|\delta(V_i)|.$$
Thus, combining inequalities, we observe
$$c(\cO) - c(\cS^{i-1}) \leq 2k |\delta(V_i)|.$$
But $|\delta(V_i)| = c(\cS^i) - c(\cS^{i-1})$ and thus
$$c(\cO) - c(\cS^{i-1}) \leq 2k (c(\cS^i) - c(\cS^{i-1}))$$
as desired.

@@

Since $\cB$ can be run in polynomial time, and $k$ is at most $|V|$, it is clear that step $2$ of $\cA$, and hence $\cA$, runs in polynomial time. At termination the algorithm returns $k$ subsets of $V$, so the algorithm returns a feasible solution.

@@

Now to see the approximation factor holds let $\cS^i$ denote the solution maintained by $\cA$ at the end of iteration $i$. Then the returned solution is $\cS^{k}$. Let $\cO$ be an optimal solution. Observe from Lemma \ref{lemma:3b1} that $c(\cS^{i}) \geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{i-1})$. We will apply this inequality inductively to achieve the bound:

@@

c(\cS^{k}) &\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{k-1}) \\
		&\geq \frac{1}{2k} c(\cO) + (1- \frac{1}{2k})(\frac{1}{2k} c(\cO) + (1- \frac{1}{2k})c(S^{k-2})) \\

@@

		&= (1-(1-\frac{1}{2k})^k)\cdot c(\cO).
